\begin{abstract}

Deep learning algorithms have tremendous potential utility in the
classification of biomedical images. For example, images acquired with retinal
optical coherence tomography (OCT) can be used to accurately classify patients
with adult macular degeneration (AMD), and distinguish them from healthy
control patients. Previous research suggests that large amounts of data are
required in order to train deep learning (DL) algorithms, because of
the large number of parameters that need to be fit. To study the data
requirements of biomedical image classfication with DL, we applied a
subsampling procedure to DL classification of AMD patients from OCT
data. We found that classification performance decreases approximately
constantly with each halving of the data. Moreover, for this task, a relatively
moderate amount of data (data from approximately 1,800 patients) may be enough
to reach performance within 5\% of the maximal performance with the full
dataset. This contrasts with sub-sampling results on CIFAR-10 data, where performance deteriorates more precipitously with These results suggest that deep learning algorithms can be trained on
moderate amounts of
data, provided that images are relatively homogenous, and the effective number
of parameters is sufficiently small. Furthermore, we demonstrate that in this
application, cross-validation with a separate test set that is not used in any
part of the training does not differ substantially from cross-validation with a
validation data-set used to determine the optimal stopping point for training.

\end{abstract}
